---
title: "Workbook"
subtitle: "Data Prescription: Introducing R for Nursing Analytics"
author: "Dr. Bret Staudt Willet"
date: "December 18, 2024"
---


# Set Up

```{r}
library(tidyverse)
library(fitbitr)
```


---


# Try it Out, Part 1: Star Wars


```{r}
starwars
```


---


```{r}
glimpse(starwars)
```


---


```{r}
starwars %>% 
  pull(name)
```


---


```{r}
starwars %>% 
  select(name, sex, height, mass)
```


---


```{r}
starwars %>% 
  filter(height <= 150) %>% 
  arrange(height)
```


---


```{r}
starwars %>% 
  select(name, starships)
```


---


```{r}
starwars %>% 
  unnest(starships) %>% 
  select(name, gender, starships)
```


---


```{r}
starwars %>% 
  unnest(starships) %>% 
  select(name, gender, starships) %>% 
  group_by(gender) %>% 
  count()
```


---


```{r, warning=FALSE}
starwars %>% 
  ggplot(aes(x = mass, y = height, color = gender)) +
  geom_point()
```


---


# Try it Out, Part 2: Fitbit Data


---


### Daily Steps

```{r, message=FALSE}
steps_df <- read_csv("data/steps2024.csv")
```

```{r}
tail(steps_df, 15)
```


---


```{r, warning=FALSE}
steps_df %>%
  ggplot(aes(x = date, y = steps)) +
  geom_point()
```


---


```{r, message=FALSE}
steps_df %>%
  ggplot(aes(x = date, y = steps)) +
  geom_point() +
  geom_smooth()
```


---


```{r, message=FALSE}
steps_df %>%
  ggplot(aes(x = date, y = steps)) +
  geom_point() +
  geom_smooth() +
  xlab(NULL) +
  ylab("Steps per Day")
```


---


```{r, message=FALSE}
steps_df %>%
  ggplot(aes(x = date, y = steps)) +
  geom_point() +
  geom_smooth() +
  xlab(NULL) +
  ylab("Steps per Day") +
  ggtitle("Number of Daily Steps")
```


---


```{r, message=FALSE}
steps_df %>%
  ggplot(aes(x = date, y = steps)) +
  geom_point() +
  geom_smooth() +
  xlab(NULL) +
  ylab("Steps per Day") +
  ggtitle("Number of Daily Steps") +
  theme_bw()
```


---


```{r, message=FALSE}
steps_df %>%
  ggplot(aes(x = date, y = steps)) +
  geom_point(color = "blue", alpha = 0.5) +
  geom_smooth(color = "green") +
  xlab(NULL) +
  ylab("Steps per Day") +
  ggtitle("Number of Daily Steps") +
  theme_bw()
```


---


```{r, message=FALSE}
steps_df %>%
  ggplot(aes(x = date, y = steps)) +
  geom_point(color = "blue", alpha = 0.5) +
  geom_smooth(color = "green") +
  xlab(NULL) +
  scale_y_continuous(labels = scales::comma) +
  ylab("Steps per Day") +
  ggtitle("Number of Daily Steps") +
  theme_bw()
```


---


### Intraday Steps

```{r, message=FALSE}
steps_intraday <- read_csv("data/intraday-steps-11-28-2024.csv")
```

```{r}
head(steps_intraday, 15)
```


---


```{r, message=FALSE}
steps_intraday %>%
  ggplot(aes(x = time, y = steps)) +
  geom_point(alpha = 0.5) +
  geom_smooth() +
  xlab(NULL) +
  ylab("Steps by Minute") +
  ggtitle("Thanksgiving Steps") +
  theme_bw()
```


---


### Heart Rate

```{r, message=FALSE}
hr <- read_csv("data/hr-11-28-2024.csv")
```

```{r}
head(hr, 15)
```


---


```{r, message=FALSE}
hr %>%
  ggplot(aes(x = time, y = heart_rate)) +
  geom_point(alpha = 0.5) +
  geom_smooth() +
  xlab(NULL) +
  ylab("Beats per Minute") +
  ggtitle("Thanksgiving Heart Rate") +
  theme_bw()
```


---


# Comparison 

```{r}
comparison_day <- 
  hr %>%
  full_join(steps_intraday, by = "time")
```

```{r, warning=FALSE}
comparison_day %>%
  ggplot(aes(x = steps, y = heart_rate)) +
  geom_point() +
  xlab("Steps per Minute") +
  ylab("Beats per Minute") +
  ggtitle("Heart Rate vs. Steps") +
  theme_bw()
```


---


# Activity Summary - Thanksgiving Day

```{r, message=FALSE}
activity_df <- read_csv("data/activity-11-28-2024.csv")
```

```{r}
glimpse(activity_df)
```


---


# Activity Summary - Trend

```{r, message=FALSE}
health_df <- read_csv("data/activity-trend-2024.csv")
```

```{r}
head(health_df, 15)
```


---


```{r, message=FALSE, warning=FALSE}
health_df %>%
  ggplot(aes(x = date, y = resting_hr)) +
  geom_point() +
  geom_smooth() +
  xlab(NULL) +
  ylab("Beats per Minute while Resting") +
  ggtitle("Resting Heart Rate") +
  theme_bw()
```


---


```{r, message=FALSE, warning=FALSE}
health_df %>% 
  ggplot(aes(x = steps, y = resting_hr)) +
  geom_point() +
  geom_smooth() +
  xlab("Daily Steps") +
  scale_x_continuous(labels = scales::comma) +
  ylab("Beats per Minute while Resting") +
  ggtitle("Resting Heart Rate vs. Daily Steps") +
  theme_bw()
```


---


```{r, message=FALSE, warning=FALSE}
health_df %>%
  ggplot(aes(x = steps, y = resting_hr)) +
  geom_point() +
  geom_smooth() +
  xlim(0, 15000) +
  scale_x_continuous(label = scales::comma) +
  xlab("Daily Steps") +
  ylim(0, 70) +
  ylab("Beats per Minute while Resting") +
  ggtitle("Resting Heart Rate vs. Daily Steps") +
  theme_bw()
```


---


# Helpful Resources and Troubleshooting


---


### Resources

**Beginners:**

- [RStudio Beginners' Guide](https://education.rstudio.com/learn/beginner/)
- Book: [*Data Science in Education Using R*](https://datascienceineducation.com)
  - [Physical copy of DSIEUR](https://www.routledge.com/Data-Science-in-Education-Using-R/Estrellado-Freer-Mostipak-Rosenberg-Velasquez/p/book/9780367422257)
  - [Even more resources from DSIEUR](https://datascienceineducation.com/c18.html)

**Intermediates:**

- [RStudio Intermediates' Guide](https://education.rstudio.com/learn/intermediate/)
- Book: [*R for Data Science*](https://r4ds.had.co.nz/)

**Experts:**

- [RStudio Experts' Guide](https://education.rstudio.com/learn/expert/)
- Book: [*Data Analysis in Medicine and Health using R*](https://bookdown.org/drki_musa/dataanalysis/)
- Book: [*Learning Statistics with R*](https://learningstatisticswithr.com/)


---


# Troubleshooting

- Try to find out what the specific problem is:
  - Identify what is *not* causing the problem
- "Unplug and plug it back in" - restart R; close and reopen R
- Seek out workshops and other learning opportunities
- Reach out to others! Sharing what is causing an issue can often help to clarify the problem
  - [Posit Community forum](https://forum.posit.co/) (highly recommended!)
  - [Contact Bret!](https://bretsw.com)
- Describe your woes to a GenAI chatbot (e.g., ChatGPT)
- General strategies on learning more: [Chapter 17 of *Data Science in Education Using R*](https://datascienceineducation.com/c17.html)



---


# Questions

- Email: [bret.staudtwillet@fsu.edu](mailto:bret.staudtwillet@fsu.edu)
- Website: [bretsw.com](https://bretsw.com)
- GitHub: [github.com/bretsw/](https://github.com/bretsw/)


---


# Materials

- Slides: [bretsw.com/nursing24-workshop](https://bretsw.com/nursing24-workshop)
- Source Materials: [github.com/bretsw/nursing24-workshop](https://github.com/bretsw/nursing24-workshop)
- Workbook: [github.com/bretsw/nursing24-workshop/blob/main/index.Rmd](https://github.com/bretsw/nursing24-workshop/blob/main/workbook.Rmd)
